raptr: Representative and Adequate Prioritisation Toolkit in R

### 2016-11-04 code/rmarkdown/article.Rmd
+ edits to ms + increase blm value

1. An underlying aim in conservation planning is to maximise the long-term persistence of biodiversity. To fulfill this aim, the processes that sustain biodiversity must be conserved. Many biodiversity processes can be preserved by conserving a representative sample of each biodiversity feature (species or ecosystem). For example, representing ability for individuals to persist in a range of climatic conditions might be crucial for the long-term persistence of a species in a world where environmental change is accelerating. Yet current approaches overwhelmingly focus on securing a target proportion of the features' geographic range.

2. To address this issue, we developed the \\texttt{raptr: Representative and Adequate Prioritisation Toolkit in R} to guide reserve selection using explicit targets for representing biodiversity processes as well as area protection targets. Users set \"space targets\" to secure a representative sample of each feature within an attribute space that reflects a biodiversity process (eg. an attribute space expressing variation in genetic characteristics between individuals, or variation in climatic conditions between areas). Users also set \"amount targets\"--similar to conventional reserve selection methods--to secure a target proportion of the feature's geographic distribtuion. We demonstrated the functionality of this package using a simulated study and two case studies. In each study, we generated solutions that secured a proportion of the species' distributions using amount targets--representing conventional reserve selection--and compared them with solutions generated using both amount and space targets.

3. We show that explicitly considering biodiversity processes in reserve selection can substantially change the solution. The simulations suggest that including representativeness in conservation planning is important where biodiversity features have multi-model distributions in an attribute space. The case studies show that explicitly considering space-based targets for biodiversity processes can return solutions that could be far more effective in achieving long term biodiversity persistence and do not require a much greater area reserved than traditional approaches.

@@

<!-- biodiversity processes are important -->
Prioritisations must capture the ecological and evolutionary processes that sustain biodiversity [@r3; @r255; @r9]. Ecological processes are important for biodiversity to persist over short time-scales. For instance, predator-prey processes can mediate ecosystem structure and function through top-down control [@r476]. Additionally, migratory processes allow many species to access suitable habitat throughout the year [@r477]. On the other hand, evolutionary processes are important for biodiversity to persist over long time-scales. Adaptive evolutionary processes foster resilience against environmental change [@r3; @r43], and neutral evolutionary processes maintain genetic diversity [@r5; @r67; @r63]. In recent decades, a wealth of data relevant to biodiversity processes has become available [eg. bioclimatic, distribution, genetic, phylogenetic, and trait data; @r406; @r33; @r460; @r453]. This is sparked significant interest in using this data to make better conservation decisions [@r68; @r9; @r473; @r247; @r448; @r468; @r2; @r407; @r61; @r470; @r475; @r469].

<!-- current methods for preserving biodiversity processes -->
Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Marxan} and \texttt{Zonation}; @r429; @r22]. The cornerstone of these tools is that decision makers can set targets for a proportion of the geographic range of species, or ecosystems (termed features). To incorporate some information on biodiversity processes, current methods involve partitioning the features into different groups as a pre-processing step [@r9; @r247; @r448]. These groups are partitioned based on a proxy variable that relates to a biodiversity process of interest [@r473]. For instance, species ranges' have been partitioned according to habitat discontinuities to capture adaptive processes [@r2]. However, by partitioning features into different groups, it is not clear how well a given solution samples the original proxy variable, and ultimately, how well the solution captures the biodiversity process of interest. To overcome this limitation, Faith and Walker [-@r218] developed a decision support tool (\texttt{DIVERSITY}) to explicitly sample proxy variables to secure biodiversity processes. However, this tool is not widely used today, in part, because it cannot accommodate explicit targets on biodiversity patterns for multiple features.

<!-- aims of this paper -->
One of the key issues in reserve selection is the lack of a unified decision support tool that uses explicit and separate targets for biodiversity patterns and processes. To begin to fill this gap, we united the ideas underpinning \texttt{DIVERSITY} and \texttt{MARXAN} into new formulations of the reserve selection problem, and implemented them in the \texttt{raptr R} package. This \texttt{R} package provides decision makers with the tools to generate prioritisations that preserve both biodiversity patterns and processes. Here, we aim to the provide an in-depth understanding of this \texttt{R} package and explore the functionality.

@@

To illustrate these concepts, consider the following example: we wish to develop a prioritisation for a single species that has four populations. However, we can only afford to preserve three of the four populations. To ensure that we preserve a representative sample, we want to want preserve populations inhabiting different environmental conditions. We obtained annual rainfall (mm) and temperature ($^{\circ}$C) data at the location of each population. We used this data to construct a two-dimensional environmental attribute space. Next, we generated demand points as equi-distant points inside this space. By comparing the distributions of demand points and populations in the attribute space, we can identify a suitable prioritisation (Fig. 1). We can see that a solution that prioritises both populations _A_ and _C_, effectively "doubles-up" on the same environmental characteristics and would contain in considerable redundancy. Instead, a more representative sample of the intra-specific variation could be captured by securing populations $A$, $B$, and $D$. This example demonstrates how attribute spaces and demand points relate to biodiversity processes, and can be used to guide reserve selection process .

The formulations used to express the reserve selection problem in the \texttt{raptr R} package are based on a combination of the \texttt{Marxan}, \texttt{DIVERSITY}, and the uncapacitated facility location problems [@r451]. Although the \texttt{raptr R} package utlises two novel formulations, for brevity, we will define the simpler formulation--referred to as the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1. These formulations are named after the unreliable and reliable facility location formulations from which they are based upon [@r16]. The key difference between these two formulations is that the reliable formulation explicitly considers the probability that planning units are occupied when calculating the proportion of an attribute space sampled in a solution. All mathematical terms defined hereafter are described in Table S1. For convenience, the cardinality of sets will be denoted using the same symbol used to denote the variable.

Define $F$ to be the set of features one wishes to conserve (indexed by $f$). Let $J$ be a set of planning units (indexed by $j$). Also, let $A_j$ denote the area, and $C_j$ denote the cost of preserving planning unit $j \in J$. To assess the extent to which each feature is secured in a given prioritisation, let $q_{fj}$ denote the probability of feature $f$ occupying planning unit $j$. The level of fragmentation associated with a prioritisation is parametrised as the net exposed boundary length. Let the shared edges between each planning unit $j \in J$ and $k \in J$ be $e_{jk}$.

Let $S$ denote a set of attribute spaces (indexed by $s$). Each $j \in J$ is associated with spatially explicit data that represent coordinates for each attribute space $s \in S$. Let $I_{fsi}$ denote a set of demand points (indexed by $i$) for each feature $f \in F$ and each attribute space $s \in S$. Let $\lambda_{fsi}$ denote the weighting for each demand point $i \in I$, $f \in F$ and $s \in S$. Let $d_{fsij}$ denote the distance between each demand point $i \in I$ and each planning unit $j \in J$ for each feature $f \in F$ and attribute space $s \in S$. To describe the inherent variation in the distribution of demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote the distance between each demand point $i \in I$ and the centroid of the demand points. Demand points with greater weight $\lambda_{fsi}$ are more important, and the optimal solution will be likely to select planning units close to highly weighted demand points. As a consequence, the decision maker will need to choose an appropriate weighting for each demand point.

Targets are used to ensure that prioritisations adequately preserve each feature. Amount-based targets are used to ensure that the total amount of habitat preserved is sufficient. Let $T_f$ denote the expected amount of area that needs to be preserved for each feature $f \in F$. Space-based targets ensure that a representative sample of each feature is secured. Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $a \in A$. For convenience, all targets are expressed as proportions in the \texttt{R} package.

The control variables for the unreliable formulation are the $B$, $T_{s}$, and $\tau_{sa}$ variables.

\begin{align*}
T_s &= \text{amount target for feature $f$} \tag*{eqn 1a}\\
%
\tau_{sa} &= \parbox{25em}{representation target for feature $f$ in attribute space $a$ according to the corresponding demand points} \tag*{eqn 1b}\\
%
B &= \text{boundary length modifier (BLM): penalise fragmented solutions} \tag*{eqn 1c}\\
\end{align*}

@@

Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is assigned to a selected planning unit $J$ where $X_j = 1$. The weighted distance between the demand point and its assigned planning unit $\lambda_{fsi} d_{fsij}$ is used to assess how well the demand point is represented in a given solution. Generally, demand points are assigned to the closest selected planning units (unless particularly low space-based targets are used).

@@

The addition of a positive boundary length modifier (BLM) parameter resulted in a well-connected solution that secured an adequate proportion (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % ) and representative sample (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$space.held,2)` % sampled) of the species geographic distribution (Fig. 3d). This solution contains `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n` planning units--a few more than the previously discussed solutions--to ensure that the solution secures a representative proportion of the species geographic distribution in a configuration that is not highly fragmented. This result suggests that the combination of space targets and boundary length modifiers may yield solutions that contain more planning units, since the solution generated using amount targets and a boundary length modifier (Fig. 3c) contained the sample number of planning units as the solution with just the amount targets (Fig. 3a). Overall, these results show that under the simplest of conditions, the reliance on just amount targets can cause an under-specified reserve selection problem that is unlikely to return a suitable solution for implementation.

@@

The normally distributed species was simulated to contain a single core area where individuals are most prevalent and marginal areas where individuals are less likely to occur (Fig. 2b). The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets contained `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$n` planning units and concentrated conservation efforts in the core area (Fig. 3e). Since all planning units have equal costs and areas, this solution contains the planning units with the highest probabilities of occupancy. Whilst this solution satisfies the adequacy objective for a prioritisation in a cost-effective manner--it fails to fulfill the representative objective for a prioritisation (`r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % distribution sampled).

The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets resulted in a solution that is both adequate and representative in terms of the uniform species distribution (Fig. 3f). This solution used `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$n` planning units to secure `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % and sample `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % of the normal species' distribution. By using amount and geographic targets in the reserve selection problem, we have obtained a solution that concentrates conservation effort in the range core and also the range margin. However, similar to the corresponding solution for the uniform species (Fig. 3b), this solution is highly fragmented.

By using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets and a boundary length modifier parameter ($BLM = 1$), we obtained a solution that fulfills adequacy, representativeness and connectivity objectives (Fig. 3h). This solution contains `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n` planning units. Similar to the corresponding solution for the uniform species (Fig. 3d), this solution contains more planning units than any other solutions for this species (cf. Figs. 3e--3g). The results for the normally distributed species suggest that the space targets can result in solutions that secure a more representative sample of the species'--even for species without significant spatial structure.

@@

The bimodally distributed species was simulated to represent a species with a highly structured distribution, and it contains two distinct populations (Fig. 2c). The solution generated using just `r sim.params.LST[[MODE]]$amount.target*100` % amount targets assigned conservation effort to just one of the two populations (Fig. 3i). While this solution secured an adequate proportion of the species' distribution (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$amount.held,2)` % secured), it did not sample a representative proportion of the range (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled). The addition of boundary length modifiers to just amount targets resulted in a solution which sampled even less of the species distribution (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount target\nand BLM')$space.held,2)` % sampled; Fig. 3k). However, the addition of geographic space targets resolved these issues.

@@

## CASE STUDY 2
### Methods
Here we used space-based targets to generate a prioritisation securing a representative sample of a species' intra-specific genetic variation. We used species occurrence and multilocus AFLP data collected by the international IntraBioDiv project in the European Alps [@r463; @r461; @r462]. Although this dataset contains multiple plant species, we used data for the `r tolower(cs2.params.LST[[MODE]]$common.name)` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}), a wide-spread species with significant genetic structure [@r462] as a simple case study. Members of the IntraBioDiv project collected data using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25 km; Fig. 7a). They visited every second grid cell, and if the `r tolower(cs2.params.LST[[MODE]]$common.name)` was detected in a cell, samples were collected from three individuals. Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r472], and used to construct matrices denoting the presence/absence of polymorphisms at loci. In total, `r structurer::n.samples(cs2.spp.StructureData.LST[[1]])` individuals were genotyped at `r structurer::n.loci(cs2.spp.StructureData.LST[[1]])` markers.

We used non-metric multi-dimensional scaling [NMDS; using Gower distances to accommodate sparsity; @r459; implemented the \texttt{cluster R} package; @r457] to ordinate the presence (or absence) of locus-specific alleles wthin individuals into two continuous variables [implemented in the \texttt{vegan R} package; @r458]. These continuous variables described the main axes of genetic variation within the species. We calculated the average of the values associated with individuals in each grid cell. These values were used to create a genetic attribute space (Figs. 7c and 7d). To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using inverse great circle distances based on the grid cells' centroids [using the \texttt{ape R} package; @r464].

@@

A two-dimensional attribute genetic space was able to capture most of the variation between individuals (stress = `r round(cs2.spp.nmds.LST[[1]]$stress,2)`; Figs. 7c--7d). Planning units located near each were contained individuals with similar genetic characteristics (Moran's I: NMDS axis 1, $I = `r round(cs1.nmds.MoransI[[1]][[1]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[1]]$p.value)`$; Moran's I: NMDS axis 2, $I = `r round(cs1.nmds.MoransI[[1]][[2]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[2]]$p.value)`$). In terms of the average genetic characteristics of individuals in the planning units, they tended to cluster into two main groups, with evidence of within-group structure inside the larger group (Fig 8c). This analysis supports previous work by Alvarez _et al._ [-@r461] who also found evidence of genetic structure within this species.

The solution generated using just the amount target failed to preserve a representative portion of the species' genetic variation (`r round(cs2.spp.DF$genetic[1],2)` % sampled; Figs. 8a and 8c). This solution only sampled planning units that contained individuals belonging to one of the main two groups. On the other hand, the solution generated using additional genetic targets selected planning units belonging to both of the main groups. Note that since these solutions were generated using opportunity cost data, the solution generated using only the amount target is essentially the minimum number of least costly areas needed to fulfill the amount target. The only difference between the two solutions is a single planning unit. By swapping this single planning unit, the solution generated using amount and genetic targets was able to preserve a representative portion of the species' genetic variation (`r round(cs2.spp.DF$genetic[2],2)` % sampled; Figs. 8b and 8d) for only a minor increase in cost (`r round(cs2.spp.DF$Score[1],2)` total cost compared to `r round(cs2.spp.DF$Score[2],2)` total cost).

# Implications and future directions
The \texttt{raptr} R package provides a unified approach to reserve selection. One of the key advantages of this package is that it is general enough to incorporate any spatial variable as an attribute space. This package can accommodate intrinsic or extrinsic variation to the feature(s). For example, adaptation processes could be secured using environmental variation or genetic variation among loci under selection. Ecological processes, such as predator-prey interactions, could be secured by capturing intra-specific trophic variation. Additionally, different attribute spaces with different targets can be specified for different species. As long as the variation can be described using Euclidean distances, the \texttt{R} package can be used to obtain a representative sample (note that data may require transformation or dimension reduction to meet this assumption). This \texttt{R} package provides decision makers with the tools to generate prioritisations that secure both biodiversity patterns and the processes that maintain biodiversity. Additionally, the package contains functionality to accommodate uncertainty in the distribution of features, and also identify suitably connected reserves. Both the simulated and case study species suggest that explicitly considering proxies for biodiversity processes can significantly alter prioritisations.

>>>>

1. An underlying aim in conservation planning is to maximise the long-term persistence of biodiversity. To fulfill this aim, the processes that sustain biodiversity must be conserved. Many biodiversity processes can be preserved by conserving a representative sample of each biodiversity feature (species or ecosystem). For example, representing ability for individuals to persist in a range of climatic conditions might be crucial for the long-term persistence of a species in a world where environmental change is accelerating. Yet current approaches overwhelmingly focus on securing a target proportion of features' geographic ranges; little attention is directed towards targeting a representative sample of them.

2. To address this issue, we developed the \\texttt{raptr: Representative and Adequate Prioritisation Toolkit in R} to guide reserve selection using explicit targets for representing biodiversity processes as well as area protection targets. Users set \"space targets\" to secure a representative sample of each feature within an attribute space that reflects a biodiversity process (eg. an attribute space expressing variation in genetic characteristics between individuals, or variation in climatic conditions between areas). Users also set \"amount targets\"--similar to conventional reserve selection methods--to secure a target proportion of the feature's geographic distribution. We demonstrated the functionality of this package using a simulated study and two case studies. In each study, we generated solutions that secured a proportion of the species' distributions using amount targets--representing conventional reserve selection--and compared them to solutions generated using both amount and space targets.

3. We show that explicitly considering biodiversity processes in reserve selection can substantially change the solution. The simulations suggest that including representativeness in conservation planning is important where biodiversity features have multimodal distributions in an attribute space. The case studies show that explicitly considering space-based targets for biodiversity processes can result in solutions that could be far more effective in achieving long term biodiversity persistence and do not require a much greater area reserved than traditional approaches.

@@

<!-- biodiversity processes are important -->
Prioritisations must capture the ecological and evolutionary processes that sustain biodiversity [@r3; @r255; @r9]. Ecological processes are important for biodiversity to persist over short time-scales. For instance, predator-prey processes can mediate ecosystem structure and function through top-down control [@r476]. Additionally, migratory processes allow many species to access suitable habitat throughout the year [@r477]. On the other hand, evolutionary processes are important for biodiversity to persist over long time-scales. Adaptive evolutionary processes foster resilience against environmental change [@r3; @r43], and neutral evolutionary processes maintain genetic diversity [@r5; @r67; @r63]. In recent decades, a wealth of data relevant to biodiversity processes has become available [eg. bioclimatic, distribution, genetic, phylogenetic, and trait data; @r33; @r460; @r478; @r453]. This has sparked significant interest in using this data to make better conservation decisions [@r68; @r9; @r473; @r247; @r448; @r468; @r2; @r407; @r61; @r470; @r475; @r469].

<!-- current methods for preserving biodiversity processes -->
Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Marxan} and \texttt{Zonation}; @r429; @r22]. The cornerstone of these tools is that they allow decision makers to set targets to obtain solutions that secure a proportion of the geographic range of biological features (populations, species, or ecosystems of conservation interest). To incorporate some information on biodiversity processes, current methods involve partitioning the features into different groups as a pre-processing step [@r9; @r247; @r448]. These groups are partitioned based on a proxy variable that relates to a biodiversity process of interest [@r473]. For instance, species ranges' have been partitioned according to habitat discontinuities to capture adaptive processes [@r2]. However, by partitioning features into different groups, it is not clear how well a given solution samples the original proxy variable, and ultimately, how well the solution captures the biodiversity process of interest. This limitation is also present in some of the earliest reserve selection methods [eg. @r295], and to overcome it, Faith and Walker [-@r218] developed a decision support tool (\texttt{DIVERSITY}) to explicitly sample proxy variables to secure biodiversity processes. However, this tool is not widely used today, in part, because it cannot accommodate explicit targets on biodiversity patterns for multiple features (unlike and \texttt{Marxan} and \texttt{Zonation}).

<!-- aims of this paper -->
One of the key issues in reserve selection is the lack of a unified decision support tool that uses explicit and separate targets for biodiversity patterns and processes. To begin to fill this gap, we united the ideas underpinning \texttt{DIVERSITY} and \texttt{Marxan} into new formulations of the reserve selection problem, and implemented them in the \texttt{raptr R} package. This \texttt{R} package provides decision makers with the tools to generate prioritisations that preserve both biodiversity patterns and processes. Here, we aim to the provide an in-depth understanding of this \texttt{R} package and explore its functionality.

@@

To illustrate these concepts, consider the following example: we wish to develop a prioritisation for a single species that has four populations. However, we can only afford to preserve three of the four populations. To ensure that we preserve a representative sample, we want to preserve populations inhabiting different environmental conditions. We obtained annual rainfall (mm) and temperature ($^{\circ}$C) data at the location of each population. We used this data to construct a two-dimensional environmental attribute space. Next, we generated demand points as equi-distant points inside this space. By comparing the distributions of demand points and populations in the attribute space, we can identify a suitable prioritisation (Fig. 1). We can see that a solution that prioritises both populations _A_ and _C_, effectively "doubles-up" on the same environmental characteristics and would contain in considerable redundancy. Instead, a more representative sample of the intra-specific variation could be captured by securing populations $A$, $B$, and $D$. This example demonstrates how attribute spaces and demand points relate to biodiversity processes, and can be used to guide the reserve selection process.

The formulations used to express the reserve selection problem in the \texttt{raptr R} package are based on a combination of the \texttt{Marxan}, \texttt{DIVERSITY}, and the uncapacitated facility location problems [@r451]. Although the \texttt{raptr R} package utlises two novel formulations, for brevity, we will define the simpler formulation--referred to as the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1. These formulations are based on unreliable and reliable facility location problems and have been named accordingly [@r16]. The key difference between these two formulations is that the reliable formulation explicitly considers the probability that planning units are occupied by features when calculating the proportion of an attribute space sampled in a solution. All mathematical terms defined hereafter are described in Table S1. For convenience, the cardinality of sets will be denoted using the same symbol used to denote the set.

Define $F$ to be the set of features one wishes to conserve (indexed by $f$). Let $J$ be a set of planning units (indexed by $j$). Also, let $A_j$ denote the area, and $C_j$ denote the cost of preserving planning unit $j \in J$. To assess the extent to which each feature is secured in a given prioritisation, let $q_{fj}$ denote the probability of feature $f$ occupying planning unit $j$. The level of fragmentation associated with a prioritisation is parametrised as the total exposed boundary length (as in \texttt{Marxan}). Let the shared edges between each planning unit $j \in J$ and $k \in J$ be $e_{jk}$.

Let $S$ denote a set of attribute spaces (indexed by $s$). Each $j \in J$ is associated with spatially explicit data that represent coordinates for each attribute space $s \in S$. Let $I_{fsi}$ denote a set of demand points (indexed by $i$) for each feature $f \in F$ and each attribute space $s \in S$. Let $\lambda_{fsi}$ denote the weighting for each demand point $i \in I$, $f \in F$ and $s \in S$. Let $d_{fsij}$ denote the distance between each demand point $i \in I$ and each planning unit $j \in J$ for each feature $f \in F$ and attribute space $s \in S$. To describe the inherent variation in the distribution of demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote the distance between each demand point $i \in I$ and the centroid of the demand points. Demand points with greater weight $\lambda_{fsi}$ are more important, and the solutions that select planning units close to highly weighted demand points will be more favorable. As a consequence, the decision maker will need to choose an appropriate weighting for each demand point.

Targets are used to ensure that prioritisations adequately preserve each feature. Amount-based targets are used to ensure that the total amount of habitat preserved is sufficient. Let $T_f$ denote the expected amount of area that needs to be preserved for each feature $f \in F$. Space-based targets ensure that a representative sample of each feature is secured. Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $s \in S$. For convenience, all targets are expressed as proportions in the \texttt{R} package.

The control variables for the unreliable formulation are the $B$, $T_{s}$, and $\tau_{fs}$ variables.

\begin{align*}
T_s &= \text{amount target for feature $f$} \tag*{eqn 1a}\\
%
\tau_{fs} &= \parbox{25em}{representation target for feature $f$ in attribute space $s$ according to the corresponding demand points} \tag*{eqn 1b}\\
%
B &= \text{boundary length modifier (BLM) to penalise fragmented solutions} \tag*{eqn 1c}\\
\end{align*}

@@

Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is assigned to a selected planning unit $J$ where $X_j = 1$. The weighted distance between the demand point and its assigned planning unit $\lambda_{fsi} d_{fsij}$ is used to assess how well the demand point is represented in a given solution. Generally, demand points are represented by their closest selected planning units (unless particularly low space-based targets are used).

@@

The addition of a positive boundary length modifier (BLM) parameter resulted in a well-connected solution that secured an adequate proportion (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % ) and representative sample (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$space.held,2)` % sampled) of the species geographic distribution (Fig. 3d). This solution contains `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n` planning units--a few more than the previously discussed solutions--to ensure that the solution secures a representative proportion of the species geographic distribution in a configuration that is not highly fragmented. This result suggests that the combination of space targets and boundary length modifiers may yield solutions that contain more planning units, since the solution generated using amount targets and a boundary length modifier (Fig. 3c) contained the same number of planning units as the solution with just the amount targets (Fig. 3a). Overall, these results show that under the simplest of conditions, the reliance on just amount targets can cause an under-specified reserve selection problem that is unlikely to return a suitable solution for implementation.

@@

The normally distributed species was simulated to contain a single core area where individuals were most prevalent, and marginal areas where individuals were less likely to occur (Fig. 2b). The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets contained `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$n` planning units and concentrated conservation efforts in the core area (Fig. 3e). Since all planning units have equal costs and areas, this solution contains the planning units with the highest probabilities of occupancy. Whilst this solution satisfies the adequacy objective for a prioritisation in a cost-effective manner--it fails to fulfill the representative objective for a prioritisation (`r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % distribution sampled).

The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets resulted in a solution that is both adequate and representative in terms of the uniform species' distribution (Fig. 3f). This solution used `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$n` planning units to secure `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % and sample `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % of the normal species' distribution. By using amount and geographic targets in the reserve selection problem, we have obtained a solution that concentrates conservation effort in the range core and also the range margin. However, similar to the corresponding solution for the uniform species (Fig. 3b), this solution is highly fragmented.

By using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets and a boundary length modifier parameter ($BLM = 1$), we obtained a solution that fulfills adequacy, representativeness and connectivity objectives (Fig. 3h). This solution contains `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n` planning units. Similar to the corresponding solution for the uniform species (Fig. 3d), this solution contains more planning units than any other solutions for this species (cf. Figs. 3e--3g). The results for the normally distributed species suggest that the space targets can result in solutions that secure a more representative sample of the species--even for species without significant spatial structure.

@@

The bimodally distributed species was simulated to represent a species with a highly structured distribution, and it contains two distinct populations (Fig. 2c). The solution generated using just `r sim.params.LST[[MODE]]$amount.target*100` % amount targets assigned conservation effort to just one of the two populations (Fig. 3i). While this solution secured an adequate proportion of the species' distribution (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$amount.held,2)` % secured), it did not sample a representative proportion of the range (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled). The addition of boundary length modifiers to amount targets resulted in a solution which sampled even less of the species distribution (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount target\nand BLM')$space.held,2)` % sampled; Fig. 3k). However, the addition of geographic space targets resolved these issues.

@@

## CASE STUDY 2
### Methods
Here we used space-based targets to generate a prioritisation securing a representative sample of a species' intra-specific genetic variation. We used species occurrence and genetic data collected by the international IntraBioDiv project in the European Alps [@r463; @r461; @r462]. Although this dataset contains multiple plant species, we used data for the `r tolower(cs2.params.LST[[MODE]]$common.name)` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}), a wide-spread species with significant genetic structure [@r462] as a simple case study. Members of the IntraBioDiv project collected data using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25 km; Fig. 7a). They visited every second grid cell, and if the `r tolower(cs2.params.LST[[MODE]]$common.name)` was detected in a cell, samples were collected from three individuals. Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r472], and used to construct matrices denoting the presence/absence of polymorphisms at loci. In total, `r structurer::n.samples(cs2.spp.StructureData.LST[[1]])` individuals were genotyped at `r structurer::n.loci(cs2.spp.StructureData.LST[[1]])` markers.

We used non-metric multi-dimensional scaling [NMDS; using Gower distances to accommodate sparsity; @r459; implemented the \texttt{cluster R} package; @r457] to ordinate the presence (or absence) of locus-specific alleles within individuals into two continuous variables [implemented in the \texttt{vegan R} package; @r458]. These continuous variables described the main axes of genetic variation within the species. We calculated the average of the values associated with individuals in each grid cell. These values were used to create a genetic attribute space (Figs. 7c and 7d). To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using inverse great circle distances based on the grid cells' centroids [using the \texttt{ape R} package; @r464].

@@

A two-dimensional attribute genetic space was able to capture most of the variation between individuals (stress = `r round(cs2.spp.nmds.LST[[1]]$stress,2)`; Figs. 7c--7d). Planning units located near each other contained individuals with similar genetic characteristics (Moran's I: NMDS axis 1, $I = `r round(cs1.nmds.MoransI[[1]][[1]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[1]]$p.value)`$; Moran's I: NMDS axis 2, $I = `r round(cs1.nmds.MoransI[[1]][[2]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[2]]$p.value)`$). In terms of the average genetic characteristics of individuals in the planning units, they tended to cluster into two main groups, with evidence of within-group structure inside the larger group (Fig 8c). This analysis supports previous work by Alvarez _et al._ [-@r461] who also found evidence of genetic structure within this species.

The solution generated using just the amount target failed to preserve a representative portion of the species' genetic variation (`r round(cs2.spp.DF$genetic[1],2)` % sampled; Figs. 8a and 8c). This solution only sampled planning units that contained individuals belonging to one of the two main groups. On the other hand, the solution generated using additional genetic targets selected planning units belonging to both of the main groups. Note that since these solutions were generated using opportunity cost data, the solution generated using only the amount target is essentially the minimum number of least costly areas needed to fulfill the amount target. The only difference between the two solutions is a single planning unit. By swapping this single planning unit, the solution generated using amount and genetic targets was able to preserve a representative portion of the species' genetic variation (`r round(cs2.spp.DF$genetic[2],2)` % sampled; Figs. 8b and 8d) for only a minor increase in cost (`r round(cs2.spp.DF$Score[1],2)` total cost compared to `r round(cs2.spp.DF$Score[2],2)` total cost).

# Implications and future directions
The \texttt{raptr} R package provides a unified approach to reserve selection. One of the key advantages of this package is that it is general enough to incorporate any spatial variable as an attribute space. This package can accommodate intrinsic or extrinsic variation to the feature(s). For example, adaptation processes could be secured using environmental variation or genetic variation. Ecological processes, such as predator-prey interactions, could be secured by capturing intra-specific trophic variation. Additionally, different attribute spaces with different targets can be specified for different species. As long as the variation can be described using Euclidean distances, the \texttt{R} package can be used to obtain a representative sample (note that data may require transformation or dimension reduction to meet this assumption). This \texttt{R} package provides decision makers with the tools to generate prioritisations that secure both biodiversity patterns and the processes that maintain biodiversity. Additionally, the package contains functionality to accommodate uncertainty in the distribution of features, and also identify suitably connected reserves. Both the simulated and case study species suggest that explicitly considering proxies for biodiversity processes can significantly alter prioritisations.

The degree to which a prioritisation truly secures a representative sample of a feature depends on the attribute spaces and distribution of demand points chosen by the decision maker. Ultimately, the space-based targets are set as a proportion based on the distribution of the demand points. As a consequence, if the decision maker uses an inappropriate set of spatial variables to construct an attribute space, or an inappropriate set of demand points , then the optimal solution will not be effective. We therefore stress that decision makers must carefully consider which biodiversity processes need to be reflected in the prioritisation, and which spatial data can be used to represent these processes. Note that maximising one process can be detrimental to another. For instance, to maximise the geographic spread of a prioritisation, reserves need to be further away from each other, yet to maximise connectivity, reserves may need to be closer together. To assist in the selection of appropriate demand points, the \texttt{R} package provides several routines for generating demand points (see the `make.DemandPoints` function). These routines use kernels to automatically generate demand points based on the distribution of features in spatial data sets [using @r231; @r450].
